---
Simulation of brain mediation RPM model - getting smoothing parameters
========================================================

In this document I simulate functional data, code the two stage least square (2SLS) procedure for estimation, and 



```{r}
#library(reshape2, warn.conflicts = FALSE)
#library(geepack, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
#library(GGally, warn.conflicts = FALSE)
library(RColorBrewer, warn.conflicts = FALSE)
library(gridExtra, warn.conflicts = FALSE)
library(refund)
library(splines)
library(MASS) #ginv for Moore-Penrose generalized inverse

mypalette   <- brewer.pal(9, "PRGn")
qualpalette <- brewer.pal(8,"Set2")
```

Loading data
=========================================
  
```{r load_data_BR6, cache=TRUE}
datos <- read.table("C:\\Users\\Yenny\\Dropbox\\2012 Research\\Mediation_Research\\Data_experiment\\FDA_data3")
                    
# Selecting Right Anterior Insula (Brain Region 6)
j <- 6
BR <- datos[(1:943)+(j-1)*943,4:26]
BR_tx <- datos[(1:943)+(j-1)*943,2]
BR_subjects <- datos[(1:943)+(j-1)*943,1]
    # Removing extreme outliers
  BR_outliers <- which(abs(as.matrix(BR)) > 30 )
    for(i in BR_outliers) if(floor(i/943)+1==24){
            BR[i%%943,23] <- NA
  		  }else{
			  BR[i%%943,floor(i/943)+1] <- NA
			  }

# Build dataframe compatible with Vadim's lecture 8 code.
BR6.refund <- with(datos[(1:943)+(6-1)*943,], {
         data.frame(ID = V1, trt = V2, pain = V3)
                  })

BR6.refund$intensity <- as.matrix(BR)
```

```{r load_data_BR19, cache=TRUE}
# Selecting Right Anterior Insula (Brain Region 6)
j <- 19
BR <- datos[(1:943)+(j-1)*943,4:26]
BR_tx <- datos[(1:943)+(j-1)*943,2]
BR_subjects <- datos[(1:943)+(j-1)*943,1]
    # Removing extreme outliers
  BR_outliers <- which(abs(as.matrix(BR)) > 30 )
    for(i in BR_outliers) if(floor(i/943)+1==24){
            BR[i%%943,23] <- NA
  		  }else{
			  BR[i%%943,floor(i/943)+1] <- NA
			  }

# Build dataframe compatible with Vadim's lecture 8 code.
BR19.refund <- with(datos[(1:943)+(6-1)*943,], {
         data.frame(ID = V1, trt = V2, pain = V3)
                  })

BR19.refund$intensity <- as.matrix(BR)
```


The structure of the data is as follows:
  
* We have a stack of 21 matrices (943 by 26) each corresponding to a different brain region.
* The first column is the subject identifier (1-20)
* Second column is temperature (0,1)
* Third column is pain rating (Continuous)
* Columns 4-26 is the time series data M(t)
* Each subject has 39 to 48 repetitions

I look at the mediated effect of the treatment through brain region 6 (RAIns - Right Anterior Insula) on brain region 19 (dACC - dorsal Anterior Cingulate Cortex) 

```{r}
  source('~/GitHub/functional_mediation/sff_Mediation.R')  
  source('~/GitHub/functional_mediation/est_se_fgam.R')  
```

```{r}
fbootstrap_ML = function(dta, index){
  dta = dta[index,]
  Y = dta[,grep("Y", names(dta))]
  X = dta[,grep("X", names(dta))]
  M = t(dta[,grep("M", names(dta))])
  
  result <- tryCatch(
    {sff_Mediation(X,Y,M,mediatorMethod="fosr2s",nbasis=nbasis,norder=norder,lambda=lambda, plot=FALSE, boot=TRUE)
     },
    error=function(cond){
      message(cond)
      return(rep(NA,(2*nrow(M)^2+5*nrow(M))))
    }
    )
  return(result)
}
```


```{r}
dta = cbind(BR19.refund[,"trt"],BR19.refund[,"intensity"],BR6.refund[,"intensity"])
dta = dta[complete.cases(dta),]

X <- dta[,1]
Y <- dta[,seq(2,ncol(BR19.refund[,"intensity"])+1)]
M <- dta[,seq(ncol(BR19.refund[,"intensity"])+2,ncol(BR19.refund[,"intensity"])+ncol(BR6.refund[,"intensity"])+1)]

nbasis = 50
norder = 6
lambda = 1e-10
```

```{r}
time1 = Sys.time()
out_full <- sff_Mediation(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=FALSE, boot=FALSE)
time2 = Sys.time()
time2 - time1

save(out_full, file="~/GitHub/functional_mediation/data_application/out_full.Rdata")
```
