Simulation of brain mediation RPM model 
========================================================

In this document I simulate functional data, code the two stage least square (2SLS) procedure for estimation, and 



```{r}
#library(reshape2, warn.conflicts = FALSE)
#library(geepack, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
#library(GGally, warn.conflicts = FALSE)
library(RColorBrewer, warn.conflicts = FALSE)
library(gridExtra, warn.conflicts = FALSE)
library(refund)
library(splines)
library(MASS) #ginv for Moore-Penrose generalized inverse

mypalette   <- brewer.pal(9, "PRGn")
qualpalette <- brewer.pal(8,"Set2")
```

Loading data
=========================================
  
```{r load_data_BR6, cache=TRUE}
datos <- read.table("C:\\Users\\Yenny\\Dropbox\\2012 Research\\Mediation_Research\\Data_experiment\\FDA_data3")
                    
# Selecting Right Anterior Insula (Brain Region 6)
j <- 6
BR <- datos[(1:943)+(j-1)*943,4:26]
BR_tx <- datos[(1:943)+(j-1)*943,2]
BR_subjects <- datos[(1:943)+(j-1)*943,1]
    # Removing extreme outliers
  BR_outliers <- which(abs(as.matrix(BR)) > 30 )
    for(i in BR_outliers) if(floor(i/943)+1==24){
            BR[i%%943,23] <- NA
			  }else{
			  BR[i%%943,floor(i/943)+1] <- NA
			  }

# Build dataframe compatible with Vadim's lecture 8 code.
BR6.refund <- with(datos[(1:943)+(6-1)*943,], {
         data.frame(ID = V1, trt = V2, pain = V3)
                  })

BR6.refund$intensity <- as.matrix(BR)
```

```{r load_data_BR19, cache=TRUE}
# Selecting Right Anterior Insula (Brain Region 6)
j <- 19
BR <- datos[(1:943)+(j-1)*943,4:26]
BR_tx <- datos[(1:943)+(j-1)*943,2]
BR_subjects <- datos[(1:943)+(j-1)*943,1]
    # Removing extreme outliers
  BR_outliers <- which(abs(as.matrix(BR)) > 30 )
    for(i in BR_outliers) if(floor(i/943)+1==24){
            BR[i%%943,23] <- NA
  		  }else{
			  BR[i%%943,floor(i/943)+1] <- NA
			  }

# Build dataframe compatible with Vadim's lecture 8 code.
BR19.refund <- with(datos[(1:943)+(6-1)*943,], {
         data.frame(ID = V1, trt = V2, pain = V3)
                  })

BR19.refund$intensity <- as.matrix(BR)
```


The structure of the data is as follows:
  
* We have a stack of 21 matrices (943 by 26) each corresponding to a different brain region.
* The first column is the subject identifier (1-20)
* Second column is temperature (0,1)
* Third column is pain rating (Continuous)
* Columns 4-26 is the time series data M(t)
* Each subject has 39 to 48 repetitions

I look at the mediated effect of the treatment through brain region 6 (RAIns - Right Anterior Insula) on brain region 19 (dACC - dorsal Anterior Cingulate Cortex) 

```{r}
  source('~/GitHub/functional_mediation/sff_Mediation.R')  
  source('~/GitHub/functional_mediation/est_se_fgam.R')  
```

```{r}
fbootstrap_ML = function(dta, index){
  dta = dta[index,]
  Y = dta[,grep("Y", names(dta))]
  X = dta[,grep("X", names(dta))]
  M = t(dta[,grep("M", names(dta))])
  
  result <- tryCatch(
    {sff_Mediation(X,Y,M,mediatorMethod="fosr2s",nbasis=nbasis,norder=norder,lambda=lambda, plot=FALSE, boot=TRUE)
     },
    error=function(cond){
      message(cond)
      return(rep(NA,(2*nrow(M)^2+5*nrow(M))))
    }
    )
  return(result)
}
```


```{r}
dta = cbind(BR19.refund[,"trt"],BR19.refund[,"intensity"],BR6.refund[,"intensity"])
dta = dta[complete.cases(dta),]

X <- dta[,1]
Y <- dta[,seq(2,ncol(BR19.refund[,"intensity"])+1)]
M <- dta[,seq(ncol(BR19.refund[,"intensity"])+2,ncol(BR19.refund[,"intensity"])+ncol(BR6.refund[,"intensity"])+1)]

nbasis = 50
norder = 6
lambda = 1e-10
```


```{r eval=FALSE}
time1 = Sys.time()
out <- sff_Mediation(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=TRUE, boot=TRUE)
time2 = Sys.time()
time2 - time1

dta2  = data.frame(X=X,M=M,Y=Y)

library(boot)
library(parallel)
library(snow)
library(doSNOW)

cl <- makeCluster(4, type="SOCK")
registerDoSNOW(cl)

time3 = Sys.time()
stat = boot(data = dta2, statistic = fbootstrap_ML,R = 100,sim = "ordinary",stype = "i", parallel="snow")
time4 = Sys.time()
time4 - time3

time3 = Sys.time()
stat2 = boot(data = dta2, statistic = fbootstrap_ML,R = 100,sim = "ordinary",stype = "i", parallel="snow")
time4 = Sys.time()
time4 - time3

time3 = Sys.time()
stat3 = boot(data = dta2, statistic = fbootstrap_ML,R = 100,sim = "ordinary",stype = "i", parallel="snow")
time4 = Sys.time()
time4 - time3

time3 = Sys.time()
stat4 = boot(data = dta2, statistic = fbootstrap_ML,R = 100,sim = "ordinary",stype = "i", parallel="snow")
time4 = Sys.time()
time4 - time3

time3 = Sys.time()
stat5 = boot(data = dta2, statistic = fbootstrap_ML,R = 100,sim = "ordinary",stype = "i", parallel="snow")
time4 = Sys.time()
time4 - time3

stopCluster(cl)

stats = list(stat=stat, stat2=stat2, stat3=stat3, stat4=stat4, stat5=stat5)
save(dta2,stats, out, file="~/GitHub/functional_mediation/data_application/boot_stats.Rdata")
```

```{r}
load(file="~/GitHub/functional_mediation/data_application/boot_stats.Rdata")
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
```


```{r}
dta2$ID <- seq(1,nrow(dta2))
dta3 <- gather(dta2, Parameter,BOLD, -c(X, ID))
dta3 <- dta3 %>% separate(Parameter, c("BR", "time"))

dta3$BR <- factor(dta3$BR, levels=c("M", "Y"), labels = c("Right Anterior Insula", " dorsal Anterior Cingulate Cortex"))

dta3$time <- factor(dta3$time, levels=paste0("V", 4:26), labels = 1:23)
dta3$time <- as.numeric(as.character(dta3$time))*1.5

dta3$Condition <- factor(dta3$X, levels=c(1,0), labels = c("High", "Low"))
```

```{r dta, fig.height=4.7,fig.width=5,dpi=100, fig.path='C:/Users/Yenny/Documents/GitHub/functional_mediation/data_application/', include=FALSE}
c = ggplot(data=dta3, aes(y = BOLD, x = time, colour=Condition))
c = c +facet_grid(Condition ~ BR , scale="fixed")

c$labels = list(x = "Time in seconds", y = "Intensity", colour = "Condition")
c = c + geom_path(data = dta3, aes(x = time, y = BOLD, group = ID, colour = Condition), size = 0.7, alpha = 0.01) + guides(colour=FALSE) + stat_smooth(size=2,level = 0.95) 
c
```

```{r dta2, fig.height=3,fig.width=5,dpi=100, fig.path='C:/Users/Yenny/Documents/GitHub/functional_mediation/data_application/', include=FALSE}
c = ggplot(data=dta3, aes(y = BOLD, x = time, colour=BR))
c = c +facet_grid(. ~ BR , scale="fixed")

c$labels = list(x = "Time in seconds", y = "Intensity", colour = "BR")
c = c + geom_path(data = dta3, aes(x = time, y = BOLD, group = ID, colour = BR), size = 0.7, alpha = 0.01) + guides(colour=FALSE) + stat_smooth(size=2,level = 0.95) 
c
```

```{r}
all_stats = rbind(stats$stat$t, stats$stat2$t, stats$stat3$t, stats$stat4$t, stats$stat5$t)

Pall = 2*apply(cbind(colMeans(all_stats < 2*matrix(stats$stat$t0, ncol=ncol(all_stats), nrow=nrow(all_stats), byrow = TRUE), na.rm = TRUE),colMeans(all_stats > 2*matrix(stats$stat$t0, ncol=ncol(all_stats), nrow=nrow(all_stats), byrow = TRUE), na.rm = TRUE)),1, min)

colnames(all_stats) <- names(Pall) <- names(out)
  
source('~/GitHub/functional_mediation/sff_vec2list.R')
source('~/GitHub/functional_mediation/lplot.R')
```

```{r}
par(mfrow=c(2,1))
sff_vec2list(out, stats="d1function", plot=TRUE, returns = FALSE)
sff_vec2list(Pall, stats="d1function", plot=TRUE, returns = FALSE)

sff_vec2list(out, stats="afunction", plot=TRUE, returns = FALSE)
sff_vec2list(Pall, stats="afunction", plot=TRUE, returns = FALSE)

sff_vec2list(out, stats="absurface", plot=TRUE, returns = FALSE)
sff_vec2list(Pall, stats="absurface", plot=TRUE, returns = FALSE)

sff_vec2list(out, stats="abfunction", plot=TRUE, returns = FALSE)
sff_vec2list(Pall, stats="abfunction", plot=TRUE, returns = FALSE)

sff_vec2list(out, stats="gfunction", plot=TRUE, returns = FALSE)
sff_vec2list(Pall, stats="gfunction", plot=TRUE, returns = FALSE)
```


```{r}
  b_s_fun = sapply(1:nrow(all_stats), function(i) sff_vec2list(all_stats[i,], stats="bsurface", surface=list(s=5,t="all")))
  b_m_fun = sapply(1:nrow(all_stats), function(i) sff_vec2list(all_stats[i,], stats="bsurface", surface=list(s=12,t="all")))
  b_l_fun = sapply(1:nrow(all_stats), function(i) sff_vec2list(all_stats[i,], stats="bsurface", surface=list(s=23,t="all")))
  
  ab_s_fun = sapply(1:nrow(all_stats), function(i) sff_vec2list(all_stats[i,], stats="absurface", surface=list(s=5,t="all")))
  ab_m_fun = sapply(1:nrow(all_stats), function(i) sff_vec2list(all_stats[i,], stats="absurface", surface=list(s=12,t="all")))
  ab_l_fun = sapply(1:nrow(all_stats), function(i) sff_vec2list(all_stats[i,], stats="absurface", surface=list(s=23,t="all")))

  b_and_ab = t(rbind(b_s_fun, b_m_fun, b_l_fun, ab_s_fun, ab_m_fun, ab_l_fun))
  colnames(b_and_ab) = c(paste("bsfunction", 1:nrow(b_s_fun), sep="_"),
                         paste("bmfunction", 1:nrow(b_s_fun), sep="_"),
                         paste("blfunction", 1:nrow(b_s_fun), sep="_"),
                         paste("absfunction", 1:nrow(b_s_fun), sep="_"),
                         paste("abmfunction", 1:nrow(b_s_fun), sep="_"),
                         paste("ablfunction", 1:nrow(b_s_fun), sep="_"))
```

```{r}
  params = do.call(rbind, strsplit(colnames(all_stats), "_"))
  
  fun_all = all_stats[,!(params[,1] %in% c("absurface", "bsurface"))]
  fun_all = cbind(fun_all, b_and_ab)

params = do.call(rbind, strsplit(colnames(fun_all), "_"))
  
all_dta = data.frame(ID=rep(1:nrow(fun_all), each=ncol(fun_all)), Parameter=params[,1], time = as.numeric(params[,2]), BOLD = c(t(fun_all)))
```


```{r}
all_tbl = tbl_df(all_dta) 

all_tbl$Parameter = mapvalues(all_tbl$Parameter, from = levels(all_tbl$Parameter), to = c("ab", "ab (t=23)", "ab (t=18)", "ab (t=7)",  "a", "b (t=23)","b (t=18)","b (t=7)", "d1","d2", "c"))

all_tbl$time <- all_tbl$time*1.5

mediator_tbl = filter(all_tbl,Parameter %in% c("a","d1"))
mediator_tbl$Parameter = factor(mediator_tbl$Parameter, levels=c("a","d1"))

outcome_tbl = filter(all_tbl,Parameter %in% c("d2", "c","b (t=7)", "b (t=18)","b (t=23)"))
outcome_tbl$Parameter = factor(outcome_tbl$Parameter, levels=c("d2", "c","b (t=7)", "b (t=18)","b (t=23)"))

indirect_tbl = filter(all_tbl,Parameter %in% c("b (t=7)","b (t=18)","b (t=23)","a", "ab (t=7)","ab (t=18)","ab (t=23)", "ab"))
indirect_tbl$Parameter = factor(indirect_tbl$Parameter, levels=c("b (t=7)","b (t=18)","b (t=23)","a", "ab (t=7)","ab (t=18)","ab (t=23)", "ab"))

small_tbl = filter(all_tbl,Parameter %in% c("a", "b (t=18)","b (t=23)", "ab", "c"))
small_tbl$Parameter = factor(small_tbl$Parameter, levels=c("a", "b (t=18)","b (t=23)", "ab", "c"))
```

```{r sff_med_model, fig.height=3.2,fig.width=5,dpi=100, fig.path='C:/Users/Yenny/Documents/GitHub/functional_mediation/data_application/', include=FALSE}
c = ggplot(data=mediator_tbl, aes(y = BOLD, x = time))
c = c +facet_grid(. ~ Parameter , scale="fixed")

c$labels = list(x = "Time in seconds", y = "Intensity", colour = "Parameter")
c = c + geom_path(data = mediator_tbl, aes(x = time, y = BOLD, group = ID, colour = Parameter), size = 0.7, alpha = 0.05) + guides(colour=FALSE)
c
```

```{r sff_outcome_model, fig.height=3.2,fig.width=11,dpi=100, fig.path='C:/Users/Yenny/Documents/GitHub/functional_mediation/data_application/', include=FALSE}
c = ggplot(data=outcome_tbl, aes(y = BOLD, x = time))
c = c +facet_grid(. ~ Parameter , scale="fixed")

c$labels = list(x = "Time in seconds", y = "Intensity", colour = "Parameter")
c = c + geom_path(data = outcome_tbl, aes(x = time, y = BOLD, group = ID, colour = Parameter), size = 0.7, alpha = 0.05) + guides(colour=FALSE)
c
```

```{r sff_indirect, fig.height=4.7,fig.width=8,dpi=100, fig.path='C:/Users/Yenny/Documents/GitHub/functional_mediation/data_application/', include=FALSE}
c = ggplot(data=indirect_tbl, aes(y = BOLD, x = time))
c = c +facet_wrap(~ Parameter, ncol=4, scale="fixed")

c$labels = list(x = "Time in seconds", y = "Intensity", colour = "Parameter")
c = c + geom_path(data = indirect_tbl, aes(x = time, y = BOLD, group = ID, colour = Parameter), size = 0.7, alpha = 0.05) + guides(colour=FALSE)
c
```


