---
title: "Simulation using splines. Original. In Gypsie"
author: "Yenny Webb-Vargas"
date: "Monday, November 17, 2014"
output: html_document
---

This implements simulation for testing my functional mediation function. 

Load the library 'refund'. It also loads the library 'fda' among many others.

```{r}
cluster = TRUE
plots   = FALSE
```

```{r}
library(refund)
library(boot)
if(cluster){
  source('/home/bst/student/ywebbvar/Github/functional_mediation/sff_Mediation.R')
  source('/home/bst/student/ywebbvar/Github/functional_mediation/est_se_fgam.R')
  source('/home/bst/student/ywebbvar/Github/functional_mediation/lplot.R')
  source('/home/bst/student/ywebbvar/Github/functional_mediation/sff_vec2list.R')
  }else{
  source('~/GitHub/functional_mediation/sff_Mediation.R') 
  source('~/GitHub/functional_mediation/est_se_fgam.R')
  source('~/GitHub/functional_mediation/lplot.R')
  source('~/GitHub/functional_mediation/sff_vec2list.R')
  }
```

Defining bootstrap function:
```{r fbootstrap}
create_functional_y = function(x, m, etai, delta2=delta2, gamma=gamma, beta=beta){
  m_mat = matrix(m, ncol=length(m), nrow=length(m), byrow=T)
  y = delta2 + gamma*x + rowSums(beta*m_mat*1/(ncol(M)-1)) + etai
  y
}

fbootstrap_ML = function(dta, index){
  dta = dta[index,]
  Y = dta[,grep("Y", names(dta))]
  X = dta[,grep("X", names(dta))]
  M = t(dta[,grep("M", names(dta))])
  
  result <- tryCatch(
    {sff_Mediation(X,Y,M,mediatorMethod="fosr2s",nbasis=nbasis,norder=norder,lambda=lambda, plot=FALSE, boot=TRUE)
     },
    error=function(cond){
      message(cond)
      return(rep(NA,(2*nrow(M)^2+5*nrow(M))))
    }
    )
  return(result)
}
```

Defining 

```{r}
sub=20
nbasis = 50
norder = 6
lambda = 1e-10

if(cluster){
  iters   = 500  
  nboot   = 1000 
}else{
  iters   = 1  
  nboot   = 100 
}

len     = 40
T_sup   = 1
timevec = seq(0,T_sup, length.out=len)
  
# Create bspline basis set
basis        = create.bspline.basis(rangeval = c(0,T_sup), nbasis = 30, norder=6)
eval.bspline = eval.basis(seq(0,1,length.out=len), basis)

# Parameters for random errors
sd_epsilon = 1
sd_eta     = 0.1
```

Create parameters: covariate and treatment effects
```{r gen_pars}
# Create delta
set.seed(2903)
delta1_coef = c(rep(0,5), runif(5,-2,0), runif(10,0,2),rep(0,10)) 
delta1      = eval.bspline%*%delta1_coef

# Create alpha
set.seed(493)
alpha_coef = c(rep(0,10), runif(10,1,4), rep(0,10)) 
alpha      = eval.bspline%*%alpha_coef

# Create delta2
set.seed(73390)
delta2_coef = c(rep(0,5), runif(5,2,5), runif(12,-2,3),rep(0,8)) 
delta2      = eval.bspline%*%delta2_coef

# Create gamma
set.seed(2939)
gamma_coef = c(rep(0,11), runif(3,3,5),runif(3,4,6),runif(3,1,4), rep(0,10)) 
gamma      = eval.bspline%*%gamma_coef

# Create beta
set.seed(84930)
beta_coef = c(rep(0,15), runif(10,6,8), runif(5,2,3)) 
beta_line = eval.bspline%*%beta_coef
 
beta = matrix(ncol=length(beta_line)*2, nrow=length(beta_line))
for(i in 1:length(beta_line)) beta[i,] = c(rep(0, (i-1)), beta_line, rep(0,length(beta_line)-i+1))
beta = beta[,(length(beta_line)+1):(length(beta_line)*2)]

if(plots){
  matplot(cbind(delta1,alpha), type="l")
  matplot(cbind(delta2,gamma,beta_line), type="l")
  image(t(beta), col  = gray((0:32)/32))
  }

alphabeta = matrix(alpha, byrow=T, ncol=length(alpha), nrow=length(alpha))*beta # In beta, the columns are 's', while the rows are 't'
alphabeta_integral  = rowSums(alphabeta)*(timevec[2]-timevec[1])  # Integral of ab-function: ab(t) = \int af(s)bf(t,s)ds dt gives ab-path

```

Setting seed:
```{r}
if(cluster) RNGkind("L'Ecuyer-CMRG")
seed = 340953
set.seed(seed)
```


### Simulation 4
In this simulation, $M$ depends on $Z$ and $X$, $Y$ only depends on $M$ and not on $X$ (there is complete mediation).

```{r}
Pall  = matrix(0, iters, (2*len^2+5*len)) #2 surfaces (abs, bf), 5 functions (abf,af,gf,d1,d2)
all   = matrix(0, iters, (2*len^2+5*len))

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = cbind(int=1,X=X)
  coeff_M  = cbind(delta1,alpha)
  colnames(coeff_M) = c("delta1","alpha")

  M    = t(tcrossprod(coeff_M, design_M))+epsilon

  eta  = matrix(rnorm(sub*len,0,sd_eta), ncol=len,nrow=(sub))

  Y    = t(sapply(1:sub, function(i) create_functional_y(x=X[i], m=M[i,], etai=eta[i,], delta2=delta2, gamma=gamma, beta=beta)))

  dta  = data.frame(X=X,M=M,Y=Y)
  
  out <- sff_Mediation(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=TRUE)
  
  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=15)
    }else{
      time1 = Sys.time()
      stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
      time2 = Sys.time()
    }

  print(time2-time1)
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE),colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE)),1, min)
  all[i,] = stat$t0
  }

colnames(Pall) = colnames(all) = colnames(stat$t) = names(stat$t0)

if(plots){
  sff_vec2list(Pall[1,], stats="bsurface", plot=TRUE, returns=FALSE)
  matplot(sapply(1:10, function(i) sff_vec2list(stat$t[i,], stats="abfunction")), type="l")
}
save(all,Pall, file=paste0("SFF_Simulation_4_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Null Simulation 1. 
In this simulation, $M$ is pure noise (and is not dependent on $X$), and $Y$ only depends on $X$.

```{r}
Pall  = matrix(0, iters, (2*len^2+5*len)) #2 surfaces (abs, bf), 5 functions (abf,af,gf,d1,d2)
all   = matrix(0, iters, (2*len^2+5*len))

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  M    = epsilon

  eta  = matrix(rnorm(sub*len,0,sd_eta), ncol=len,nrow=(sub))

  Y    = t(sapply(1:sub, function(i) create_functional_y(x=X[i], m=M[i,], etai=eta[i,], delta2=delta2, gamma=gamma, beta=matrix(0, ncol=ncol(beta), nrow=nrow(beta)))))

 dta  = data.frame(X=X,M=M,Y=Y)
  
  out  = sff_Mediation(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=TRUE)
  
  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=15)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }
                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE),colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE)),1, min)
  all[i,] = stat$t0
  }

colnames(Pall) = colnames(all) = colnames(stat$t) = names(stat$t0)

if(plots){
  sff_vec2list(Pall[1,], stats="bsurface", plot=TRUE, returns=FALSE)
  matplot(sapply(1:10, function(i) sff_vec2list(stat$t[i,], stats="abfunction")), type="l")
}

save(all,Pall, file=paste0("SFF_Simulation_1_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```



### Null Simulation 2. 
In this simulation, $M$ depends on $X$, but $Y$ only depends on $X$ and not on $M$.

```{r}
Pall  = matrix(0, iters, (2*len^2+5*len)) #2 surfaces (abs, bf), 5 functions (abf,af,gf,d1,d2)
all   = matrix(0, iters, (2*len^2+5*len))

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = cbind(int=1,X=X)
  coeff_M  = cbind(delta1,alpha)
  colnames(coeff_M) = c("delta1","alpha")

  M    = t(tcrossprod(coeff_M, design_M))+epsilon

  eta  = matrix(rnorm(sub*len,0,sd_eta), ncol=len,nrow=(sub))

  Y    = t(sapply(1:sub, function(i) create_functional_y(x=X[i], m=M[i,], etai=eta[i,], delta2=delta2, gamma=gamma, beta=matrix(0, ncol=ncol(beta), nrow=nrow(beta)))))

  dta  = data.frame(X=X,M=M,Y=Y)
 
  out <- sff_Mediation(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=TRUE) 

  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=15)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }

                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE),colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE)),1, min)
  all[i,] = stat$t0
  }

colnames(Pall) = colnames(all) = colnames(stat$t) = names(stat$t0)

if(plots){
  sff_vec2list(Pall[1,], stats="bsurface", plot=TRUE, returns=FALSE)
  matplot(sapply(1:10, function(i) sff_vec2list(stat$t[i,], stats="abfunction")), type="l")
}

save(all,Pall, file=paste0("SFF_Simulation_2_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Null simulation 3
In this simulation, $M$ depends on $\delta_1$ for all subjects (and does not depend on $X$), $Y$ only depends on $M$ and not on $X$.

```{r}
Pall  = matrix(0, iters, (2*len^2+5*len)) #2 surfaces (abs, bf), 5 functions (abf,af,gf,d1,d2)
all   = matrix(0, iters, (2*len^2+5*len))

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = rep(1,sub)
  coeff_M  = cbind(delta1)
  colnames(coeff_M) = c("delta1")

  M    = t(tcrossprod(coeff_M, design_M))+epsilon

  eta  = matrix(rnorm(sub*len,0,sd_eta), ncol=len,nrow=(sub))

  Y    = t(sapply(1:sub, function(i) create_functional_y(x=X[i], m=M[i,], etai=eta[i,], delta2=delta2, gamma=rep(0, length(gamma)), beta=beta)))

  dta  = data.frame(X=X,M=M,Y=Y)
  
  out <- sff_Mediation(X,Y,t(M),mediatorMethod="fosr2s",nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=TRUE)

  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=15)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }

                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE),colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE), na.rm = TRUE)),1, min)
  all[i,] = stat$t0
  }

colnames(Pall) = colnames(all) = colnames(stat$t) = names(stat$t0)

if(plots){
  sff_vec2list(Pall[1,], stats="bsurface", plot=TRUE, returns=FALSE)
  matplot(sapply(1:10, function(i) sff_vec2list(stat$t[i,], stats="abfunction")), type="l")
}

save(all,Pall, file=paste0("SFF_Simulation_3_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```