---
title: "Simulation using splines. Original. In Gypsie"
author: "Yenny Webb-Vargas"
date: "Wednesday, September 17, 2014"
output: html_document
---

This implements simulation for testing my functional mediation function. 

Load the library 'refund'. It also loads the library 'fda' among many others.

```{r}
cluster = TRUE
plots   = FALSE
```

```{r}
library(refund)
library(boot)
if(cluster){
  source('/home/bst/student/ywebbvar/Mediation/Functional_Mediation/fMediation_ML.R')
  }else{
  source('~/GitHub/functional_mediation/fMediation_ML.R')  
  }
```

Defining bootstrap function:
```{r fbootstrap}
fbootstrap_ML <- function(dta, index){
  dta = dta[index,]
  Y = dta[,"Y"]
  X = dta[,"X"]
  M = t(dta[,seq(2+1,ncol(dta))])
  
  result = fMediation_ML(X,Y,M,mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=FALSE, boot=TRUE)
  return(result)
}
```

```{r}
sub=20
nbasis = 50
norder = 6
lambda = 1e-8

if(cluster){
  iters   = 500  
  nboot   = 1000 
}else{
  iters   = 1  
  nboot   = 50 
}

len     = 50
T_sup   = 1
timevec = seq(0,T_sup, length.out=len)
  
# Create bspline basis set
basis        = create.bspline.basis(rangeval = c(0,T_sup), nbasis = 30, norder=6)
eval.bspline = eval.basis(seq(0,1,length.out=len), basis)

# Parameters for random errors
sd_epsilon = 1
sd_eta     = 1
```

Create parameters: covariate and treatment effects
```{r gen_pars}
# Create delta
set.seed(2903)
delta1_coef = c(rep(0,5), runif(5,-2,0), runif(10,0,2),rep(0,10)) 
delta1      = eval.bspline%*%delta1_coef

# Create alpha
set.seed(493)
alpha_coef = c(rep(0,10), runif(10,1,4), rep(0,10)) 
alpha      = eval.bspline%*%alpha_coef

# Create beta
set.seed(84930)
beta_coef = c(rep(0,15), runif(5,3,10), rep(0,10)) 
beta      = eval.bspline%*%beta_coef

delta2 = 5
gamma  = 2

if(plots) matplot(cbind(delta1,alpha,beta), type="l")
```

Setting seed:
```{r}
seed = 632020
set.seed(seed)
```


### Simulation 4
In this simulation, $M$ depends on $Z$ and $X$, $Y$ only depends on $M$ and not on $X$ (there is complete mediation).

```{r}
Pall  = matrix(0, iters, len)
Pall2 = matrix(0, iters, len)
Flag  = rep(0, iters)
A     = matrix(0, iters, len)
B     = matrix(0, iters, len)
CP    = rep(0, iters)
d2    = rep(0, iters)
ptime = list()

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = cbind(int=1,X=X)
  coeff_M  = cbind(delta1,alpha)
  colnames(coeff_M) = c("delta1","alpha")

  M    = t(tcrossprod(coeff_M, design_M))+epsilon

  eta      <- rnorm(sub, 0, sd_eta) 

  design_M_Y <- cbind(int=1,X=X,M=M)
  coeff_M_Y  <- cbind(delta2,gamma, t(beta)*1/(ncol(M)-1))

  Y    <- t(tcrossprod(coeff_M_Y, design_M_Y)) + eta
  Y    <- as.numeric(Y)

  dta  = cbind('Y'=Y,'X'=X,'M' = M)
  
  out  = fMediation_ML(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=FALSE)
  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }

                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  CP[i] = out$cp
  d2[i] = out$Y_intercept
  
  if(plots)matplot(t(stat$t), type="l", col = rgb(200,1,1, 30, maxColorValue = 255))
  }

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation_4_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Null Simulation 1. 
In this simulation, $M$ is pure noise (and is not dependent on $X$), and $Y$ only depends on $X$.

```{r}
Pall  = matrix(0, iters, len)
Pall2 = matrix(0, iters, len)
Flag  = rep(0, iters)
A     = matrix(0, iters, len)
B     = matrix(0, iters, len)
CP    = rep(0, iters)
d2    = rep(0, iters)
ptime = list()

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  M    = epsilon

  eta      <- rnorm(sub, 0, sd_eta) 

  design_M_Y <- cbind(int=1,X=X)
  coeff_M_Y  <- cbind(delta2,gamma)

  Y    <- t(tcrossprod(coeff_M_Y, design_M_Y)) + eta
  Y    <- as.numeric(Y)

  dta  = cbind('Y'=Y,'X'=X,'M' = M)
  
  out  = fMediation_ML(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=FALSE)
  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }

                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  CP[i] = out$cp
  d2[i] = out$Y_intercept
  
  if(plots)matplot(t(stat$t), type="l", col = rgb(200,1,1, 30, maxColorValue = 255))
  }

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation_1_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```



### Null Simulation 2. 
In this simulation, $M$ depends on $X$, but $Y$ only depends on $X$ and not on $M$.

```{r}
Pall  = matrix(0, iters, len)
Pall2 = matrix(0, iters, len)
Flag  = rep(0, iters)
A     = matrix(0, iters, len)
B     = matrix(0, iters, len)
CP    = rep(0, iters)
d2    = rep(0, iters)
ptime = list()

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = cbind(int=1,X=X)
  coeff_M  = cbind(delta1,alpha)
  colnames(coeff_M) = c("delta1","alpha")

  M    = t(tcrossprod(coeff_M, design_M))+epsilon

  eta      <- rnorm(sub, 0, sd_eta) 

  design_M_Y <- cbind(int=1,X=X)
  coeff_M_Y  <- cbind(delta2,gamma)

  Y    <- t(tcrossprod(coeff_M_Y, design_M_Y)) + eta
  Y    <- as.numeric(Y)

  dta  = cbind('Y'=Y,'X'=X,'M' = M)
  
  out  = fMediation_ML(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=FALSE)
  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }

                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  CP[i] = out$cp
  d2[i] = out$Y_intercept
  
  if(plots)matplot(t(stat$t), type="l", col = rgb(200,1,1, 30, maxColorValue = 255))
  }

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation_2_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Null simulation 3
In this simulation, $M$ depends on $\delta_1$ for all subjects (and does not depend on $X$), $Y$ only depends on $M$ and not on $X$.

```{r}
Pall  = matrix(0, iters, len)
Pall2 = matrix(0, iters, len)
Flag  = rep(0, iters)
A     = matrix(0, iters, len)
B     = matrix(0, iters, len)
CP    = rep(0, iters)
d2    = rep(0, iters)

for(i in 1:iters){
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = rep(1,sub)
  coeff_M  = cbind(delta1)
  colnames(coeff_M) = c("delta1")

  M    = t(tcrossprod(coeff_M, design_M))+epsilon

  eta      <- rnorm(sub, 0, sd_eta) 

  design_M_Y <- cbind(int=1,X=X,M=M)
  coeff_M_Y  <- cbind(delta2,0, t(beta)*1/(ncol(M)-1))

  Y    <- t(tcrossprod(coeff_M_Y, design_M_Y)) + eta
  Y    <- as.numeric(Y)

  dta  = cbind('Y'=Y,'X'=X,'M' = M)
  
  out  = fMediation_ML(X,Y,t(M),mediatorMethod="fosr2s", nbasis=nbasis,norder=norder,lambda=lambda, plot=plots, boot=FALSE)
  if(cluster==TRUE){
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
    }else{
    stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
    }

                
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  CP[i] = out$cp
  d2[i] = out$Y_intercept
  
  if(plots)matplot(t(stat$t), type="l", col = rgb(200,1,1, 30, maxColorValue = 255))
  }

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation_3_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```