---
title: "Simulation using splines. Original. In Gypsie"
author: "Yenny Webb-Vargas"
date: "Wednesday, September 17, 2014"
output: html_document
---

This implements simulation for testing my functional mediation function. 

Load the library 'refund'. It also loads the library 'fda' among many others.
```{r}
library(refund)
library(boot)
source('/home/bst/student/ywebbvar/Mediation/Functional_Mediation/fMediation_ML.R')
```

Defining bootstrap function:
```{r fbootstrap}
fbootstrap_ML <- function(dta, index){
  dta = dta[index,]
  Y = dta[,"Y"]
  X = dta[,"X"]
  M = t(dta[,seq(2+1,ncol(dta))])
  
  result = fMediation_ML(X,Y,M,mediatorMethod="fosr2s", nbasis=50,norder=6,lambda=1e-8, plot=FALSE, boot=TRUE)
  return(result)
}
```

```{r}
sub=20
nbasis = 50
norder = 6

iters   = 1  #500  #1 
nboot   = 50 #1000 #50

len     = 50
T_sup   = 1
timevec = seq(0,T_sup, length.out=len)
  
# Create bspline basis set
basis        = create.bspline.basis(rangeval = c(0,T_sup), nbasis = 30, norder=6)
eval.bspline = eval.basis(seq(0,1,length.out=len), basis)
```

Create parameters: covariate and treatment effects
```{r gen_pars}
# Create delta
set.seed(2903)
delta1_coef = c(rep(0,5), runif(5,-2,0), runif(10,0,2),rep(0,10)) 
delta1      = eval.bspline%*%delta1_coef

# Create alpha
set.seed(493)
alpha_coef = c(rep(0,10), runif(10,1,4), rep(0,10)) 
alpha      = eval.bspline%*%alpha_coef

# Create beta
set.seed(84930)
beta_coef = c(rep(0,15), runif(5,3,10), rep(0,10)) 
beta      = eval.bspline%*%beta_coef
```

Setting seed:
```{r}
seed = 794320
```

### Null Simulation 1. 
In this simulation, $M$ is pure noise (is not dependent on $X$), and $Y$ only depends on $X$.

```{r}
Pall  = matrix(0, iters, length(Z))
Pall2 = matrix(0, iters, length(Z))
Flag  = rep(0, iters)
A     = matrix(0, iters, length(Z))
B     = matrix(0, iters, length(Z))
CP   = rep(0, iters)

set.seed(seed)
for(i in 1:iters){
  X = rep(0,sub)
  X[sample(1:sub)[1:(floor(sub/2))]] = 2
  X = X - 1 
  
  M = sapply(1:sub, function(i) rnorm(n = length(Z),mean=0, sd=1))
  Y = X + rnorm(sub, mean=0, sd=1)
  
  # I used this code to get data to check the fMediation_ML function, 
  # it was causing problems with lambdas that are not very close to 
  # zero:
  #write.csv(cbind(Y,X,t(M)), file=paste0("sim_",seed, ".csv"))
  
  dta  = cbind('Y'=Y,'X'=X,'M' =t(M))
  
  out  = fMediation_ML(X,Y,M,nbasis,norder,lambda=1e-8)
  #stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i")
  stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
  
              
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  C[i]  = out$c
  CP[i] = out$cp
  }

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation1_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Null Simulation 2. 
In this simulation, $M$ depends on $X$, but $Y$ only depends on $X$ and not on $M$.

```{r}
Pall  = matrix(0, iters, length(Z))
Pall2 = matrix(0, iters, length(Z))
Flag  = rep(0, iters)
A     = matrix(0, iters, length(Z))
B     = matrix(0, iters, length(Z))
C     = rep(0, iters)
CP   = rep(0, iters)

set.seed(seed+10)
for(i in 1:iters){ 
  X = rep(0,sub)
  X[sample(1:sub)[1:(floor(sub/2))]] = 2
  X = X - 1 
  
  M = sapply(1:sub, function(i) rnorm(n = length(Z),mean=0, sd=1)) + t((X+1)%*%t(Z))
  Y = X + rnorm(sub, mean=0, sd=1)
  
  dta  = cbind('Y'=Y,'X'=X,'M' =t(M))
  
  out  = fMediation_ML(X,Y,M,nbasis,norder,lambda=1e-8)
  #stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i") 
  stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
  
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  C[i]  = out$c
  CP[i] = out$cp
}

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation2_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Null simulation 3
In this simulation, $M$ depends on $Z$ for all subjects (and does not depend on $X$), $Y$ only depends on $M$ and not on $X$.

```{r}
Pall  = matrix(0, iters, length(Z))
Pall2 = matrix(0, iters, length(Z))
Flag  = rep(0, iters)
A     = matrix(0, iters, length(Z))
B     = matrix(0, iters, length(Z))
C     = rep(0, iters)
CP   = rep(0, iters)

set.seed(seed+20)
for(i in 1:iters){ 
  X = rep(0,sub)
  X[sample(1:sub)[1:(floor(sub/2))]] = 2
  X = X - 1 
  
  M = sapply(1:sub, function(i) rnorm(n = length(Z),mean=0, sd=1)) + Z
  Y = apply(M,2,function(m) sum(m[floor(nrow(M)/3):ceiling(nrow(M)*2/3)])) + rnorm(sub, mean=0, sd=1)
  
  dta  = cbind('Y'=Y,'X'=X,'M' =t(M))
  
  out  = fMediation_ML(X,Y,M,nbasis,norder,lambda=1e-8)
  #stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i") 
  stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
  
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  C[i]  = out$c
  CP[i] = out$cp
}
save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation3_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Simulation 4
In this simulation, $M$ depends on $Z$ and $X$, $Y$ only depends on $M$ and not on $X$ (there is complete mediation).

```{r}
P     = rep(0,iters)
Pall  = matrix(0, iters, length(Z))
Pall2 = matrix(0, iters, length(Z))
Flag  = rep(0, iters)
A     = matrix(0, iters, length(Z))
B     = matrix(0, iters, length(Z))
C     = rep(0, iters)
CP    = rep(0, iters)

set.seed(seed+30)
for(i in 1:iters){ 
  X = rep(0,sub)
  X[sample(1:sub)[1:(floor(sub/2))]] = 2
  X = X - 1 
  
  M = sapply(1:sub, function(i) rnorm(n = length(Z),mean=0, sd=1)) + t((X+1)%*%t(Z))
  Y = apply(M,2,function(m) sum(m[floor(nrow(M)/3):ceiling(nrow(M)*2/3)])) + rnorm(sub, mean=0, sd=1)
  
  dta  = cbind('Y'=Y,'X'=X,'M' =t(M))
  
  out  = fMediation_ML(X,Y,M,nbasis,norder,lambda=1e-8)
  #stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i") 
  stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
  
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  C[i]  = out$c
  CP[i] = out$cp
}
save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation4_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```


### Simulation 5. Confounding
In this simulation, $M$ depends on $Z$, $X$ and $\epsilon$, $Y$ only depends on $M$ and $\epsilon$.

```{r eval=FALSE}
P     = rep(0,iters)
Pall  = matrix(0, iters, length(Z))
Pall2 = matrix(0, iters, length(Z))
Flag  = rep(0, iters)
A     = matrix(0, iters, length(Z))
B     = matrix(0, iters, length(Z))
C     = rep(0, iters)
CP    = rep(0, iters)

set.seed(seed+40)
for(i in 1:iters){ 
  X = rep(0,sub)
  X[sample(1:sub)[1:(floor(sub/2))]] = 2
  X = X - 1 
  
  eps = sapply(1:sub, function(i) rnorm(n = length(Z),mean=0, sd=1))
  M   = t((X)%*%t(Z)) + eps
  eta = apply(eps,2,function(e) sum(e[floor(nrow(M)/3):ceiling(nrow(M)*2/3)])) + rnorm(sub, mean=0, sd=1)
  Y   = apply(M,2,function(m) sum(m[floor(nrow(M)/3):ceiling(nrow(M)*2/3)])) + eta
  
  dta  = cbind('Y'=Y,'X'=X,'M' =t(M))
  
  out  = fMediation_ML(X,Y,M,nbasis,norder,lambda=1e-8)
  #stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i") 
  stat = boot(data = dta, statistic = fbootstrap_ML,R = nboot,sim = "ordinary",stype = "i",parallel = "multicore",ncpus=10)
  
  
  # Calculating p-value
  Pall[i,] = 2*apply(cbind(colMeans(stat$t < 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                           colMeans(stat$t > 2*matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))),1, min)
  Pall2[i,] = 2*apply(cbind(colMeans(stat$t < matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE)),
                            colMeans(stat$t > matrix(stat$t0, ncol=ncol(stat$t), nrow=nrow(stat$t), byrow = TRUE))), 1, min)
  A[i,] = out$afunction
  B[i,] = out$bfunction
  C[i]  = out$c
  CP[i] = out$cp
}

save(A,B,C,CP,Pall,Pall2, file=paste0("Simulation5_",format(Sys.time(), "%Y%m%d-%H%M"),".Rdata"))
```
