---
title: "Simulation using splines. Using a constant baseline function for Mediator"
author: "Yenny Webb-Vargas"
date: "Friday, October 31, 2014"
output: html_document
---

In this document, I set up the simulation I use for the paper 'Functional Mediation Analysis'

```{r}
library(fda)
library(refund)
source('~/GitHub/functional_mediation/fMediation_ML.R')
```

```{r}
len     = 50
T_sup   = 1
timevec = seq(0,T_sup, length.out=len)
  
# Create bspline basis set
basis        = create.bspline.basis(rangeval = c(0,T_sup), nbasis = 30, norder=6)
eval.bspline = eval.basis(seq(0,1,length.out=len), basis)
```

## 1. Create parameters: covariate and treatment effects
```{r gen_pars}
# Create delta
set.seed(2903)
delta1_coef = c(rep(0,5), runif(5,-2,0), runif(10,0,2),rep(0,10)) 
delta1      = rep(0,len) # No baseline function for mediator

# Create alpha
set.seed(493)
alpha_coef = c(rep(0,10), runif(10,1,4), rep(0,10)) 
alpha      = eval.bspline%*%alpha_coef

# Create beta
set.seed(84930)
beta_coef = c(rep(0,15), runif(5,3,10), rep(0,10)) 
beta      = eval.bspline%*%beta_coef

matplot(cbind(delta1, alpha, beta), type="l")
```

## 2. Generate mediators
```{r gen_data}

set.seed(3201023)
n          = 100
sd_epsilon = 1
Z          = rbinom(n,1,0.5)
epsilon    = matrix(rnorm(n*len,0,sd_epsilon), ncol=len,nrow=(n))

design_M = cbind(int=1,Z=Z)
coeff_M  = cbind(delta1,alpha)
colnames(coeff_M) = c("delta1","alpha")

M        = t(tcrossprod(coeff_M, design_M))
M_obs    = M+epsilon

plot(M_obs[1,], xlim = c(1,ncol(M_obs)), ylim = range(M_obs, na.rm=TRUE), main = "Mediator Functions \n with Errors", ylab = "Intensity", col = rgb((200*Z[1]+1),1,1, 30, maxColorValue = 255), type = "l")
for (i in 2:nrow(M)){ 
  lines(M_obs[i,], col=rgb((200*Z[i]+1),1,1, 30, maxColorValue=255))
}
lines(beta, col=rgb(0,0,255, 250, maxColorValue=255))
legend("topright", c("beta", "treated", "control"), col=c(rgb(0,0,255, 250, maxColorValue=255), rgb((201),1,1, 255, maxColorValue=255), rgb(1,1,1, 255, maxColorValue=255)), lty=1)

hist(crossprod(t(M_obs),beta*1/(ncol(M_obs)-1)), main="Contributions from each mediator\n to Y_obs")
```

## 3. Generate outcome.
No confounding scenario:
```{r }
delta2 = 5
gamma  = 2
sd_eta = 1

eta      <- rnorm(n, 0, sd_eta) 

design_M_Y <- cbind(int=1,Z=Z,M_obs=M_obs)
coeff_M_Y  <- cbind(delta2,gamma, t(beta)*1/(ncol(M_obs)-1))

Y_obs    <- t(tcrossprod(coeff_M_Y, design_M_Y)) + eta
Y_obs    <- as.numeric(Y_obs)

hist(Y_obs)
boxplot(Y_obs~Z, main="Boxplot of Y_obs by trt")
```

# Fitting mediation models

## Using `fRegress` for both mediator and outcome model
```{r fRegress}
out  = fMediation_ML(Z,Y_obs,t(M_obs),mediatorMethod="fRegress", nbasis=50,norder=6,lambda=1e-8, plot=TRUE)

par(mfrow=c(3,1))
plot(out$afunction, type="l")  
lines(alpha, col="blue")
plot(out$bfunction, type="l")  
lines(beta, col="blue")
plot(out$abfunction, type="l")  
lines(alpha*beta, col="blue")
```

## Using `fosr2s` for mediator model, and `fRegress` for outcome model
```{r fosr2s}
out2  = fMediation_ML(Z,Y_obs,t(M_obs),mediatorMethod="fosr2s", nbasis=50,norder=6,lambda=1e-8, plot=TRUE)

par(mfrow=c(3,1))
plot(out2$afunction, type="l")  
lines(alpha, col="blue")
plot(out2$bfunction, type="l")  
lines(beta, col="blue")
plot(out2$abfunction, type="l")  
lines(alpha*beta, col="blue")
```

## Using `pfr` with smooth option `face` in outcome model
```{r pfr}
# Penalized functional regression
fit.face <- pfr(Y=Y_obs, funcs=list(M_obs), covariates=cbind(1, Z=Z),kb=40, nbasis=26, smooth.option="fpca.face", pve=0.99, method="REML")
matplot(cbind(fit.face$BetaHat[[1]], fit.face$Bounds[[1]]),
  type = 'l', lty = c(1,2,2), col = c(1,2,2), ylab = "BetaHat")
lines(beta*1/(ncol(M_obs)-1), col="blue")
```

## Using `pfr` with smooth option `sc` in outcome model
```{r}
fit.sc <- pfr(Y=Y_obs, funcs=M_obs, covariates=cbind(1,Z=Z),kb=50, nbasis=20,smooth.option="fpca.sc", pve=0.99, method="REML")

matplot(cbind(fit.sc$BetaHat[[1]], fit.sc$Bounds[[1]]),
  type = 'l', lty = c(1,2,2), col = c(1,2,2), ylab = "BetaHat")
lines(beta*1/(ncol(M_obs)-1), col="blue")
```

## Using `fgam` in outcome model
```{r fgam}
# Another take on penalized functional regression
require(mgcv)
# Fits a smooth object with 50 basis and normal cubic spline penalty with 2nd derivatives
fit <- fgam(Y_obs ~ Z + lf(M_obs,splinepars=list(k=50,m=c(3,2))))
predictions <- predict(fit)
plot(fit)
lines(seq(0,1,length=len),beta, col="blue")

newdata <- data.frame(M_obs.tmat = seq(0,1,length=len), L.M_obs = seq(1,1,length=len))
sm <- fit$smooth[[1]]
plot(PredictMat(sm, newdata)%*%fit$coef[-(1:2)], type="l")
lines(beta, col="blue")
```

# Comparison
## Plot comparing four methods for outcome model
```{r final_plot}
par(mfrow=c(2,2))
plot(out$bfunction, type="l", main="fRegress")  
lines(beta, col="blue")

plot(PredictMat(sm, newdata)%*%fit$coef[-(1:2)], type="l", main="fgam")
lines(beta, col="blue")

matplot(cbind(fit.face$BetaHat[[1]], fit.face$Bounds[[1]]),
  type = 'l', lty = c(1,2,2), col = c(1,2,2), ylab = "BetaHat", main="face")
lines(beta*1/(ncol(M_obs)-1), col="blue")

matplot(cbind(fit.sc$BetaHat[[1]], fit.sc$Bounds[[1]]),
  type = 'l', lty = c(1,2,2), col = c(1,2,2), ylab = "BetaHat", main="SC")
lines(beta*1/(ncol(M_obs)-1), col="blue")
```

##Estimates of the direct effect and intercept in outcome model
```{r final_estimates}
rbind(hat_gamma  = cbind('true'= gamma, 'fMediation_ML' = out$cp, 
                         'fgam' = fit$coef["Z"], 
                         'pfr.face' = fit.face$beta.covariates["XZ"], 
                         'pfr.sc' = fit.sc$beta.covariates["XZ"]),  
      hat_delta2 = cbind('true' = delta2, 'fMediation_ML' = out$Y_intercept, 
                         'fgam' = fit$coef["(Intercept)"], 
                         'pfr.face' = fit.face$beta.covariates[2], 
                         'pfr.sc' = fit.sc$beta.covariates[1])
      )
```

