---
title: "Troubleshoot Mediator Model"
author: "Yenny Webb-Vargas"
output: html_document
---

This is code to troubleshoot the mediator model, which is a model with one scalar covariate and one functional covariate. 

```{r}
library(refund)
library(mgcv)
```

# Varying study size

## Small study
```{r}
sub=20

# Parameters for random errors
sd_epsilon = 1 #0.01 
sd_eta     = 1 #0.01 
```

```{r simulation}
len     = 50
T_sup   = 1
timevec = seq(0,T_sup, length.out=len)
  
# Create bspline basis set
basis        = create.bspline.basis(rangeval = c(0,T_sup), nbasis = 30, norder=6)
eval.bspline = eval.basis(seq(0,1,length.out=len), basis)

# Create delta
set.seed(2903)
delta1_coef = c(rep(0,5), runif(5,-2,0), runif(10,0,2),rep(0,10)) 
delta1      = eval.bspline%*%delta1_coef

# Create alpha
set.seed(493)
alpha_coef = c(rep(0,10), runif(10,1,4), rep(0,10)) 
alpha      = eval.bspline%*%alpha_coef

# Create beta
set.seed(84930)
beta_coef = c(rep(0,15), runif(5,3,10), rep(0,10)) 
beta      = eval.bspline%*%beta_coef

delta2 = 5
gamma  = 2

matplot(cbind(delta1,alpha,beta), type="l")

# Set seed
seed = 632020
set.seed(seed)


# Generate data
  X          = rbinom(sub,1,0.5)
  
  epsilon    = matrix(rnorm(sub*len,0,sd_epsilon), ncol=len,nrow=(sub))

  design_M = cbind(int=1,X=X)
  coeff_M  = cbind(delta1,alpha)

  colnames(coeff_M) = c("delta1","alpha")

  M   = t(tcrossprod(coeff_M, design_M))+epsilon

  eta = rnorm(sub, 0, sd_eta) 

  design_M_Y = cbind(int=1,X=X,M=M)
  coeff_M_Y  = cbind(delta2,gamma, t(beta)*1/(ncol(M)-1))

  Y    = t(tcrossprod(coeff_M_Y, design_M_Y)) + eta
  Y    = as.numeric(Y)

  dta  = cbind('Y'=Y,'X'=X,'M' = M)

# Estimate functional parameter
fit <- fgam(Y ~ X + lf(M,splinepars=list(k=ifelse(sub < 52, sub-2, 50),m=c(3,2))))
predictions <- predict(fit)
plot(fit)
lines(seq(0,1,length=length(beta)), beta, col="blue")
```


## Medium study

```{r}
sub=50

# Parameters for random errors
sd_epsilon = 1 #0.01 
sd_eta     = 1 #0.01 
```

```{r sim_medium, ref.label='simulation'}
```

## Large study

```{r}
sub=100

# Parameters for random errors
sd_epsilon = 1 #0.01 
sd_eta     = 1 #0.01 
```

```{r sim_large, ref.label='simulation'}
```


## Very large study

```{r}
sub=1000

# Parameters for random errors
sd_epsilon = 1 #0.01 
sd_eta     = 1 #0.01 
```

```{r sim_verylarge, ref.label='simulation'}
```

# Varying error in outcome model


## Unit error

```{r}
sub=20

# Parameters for random errors
sd_epsilon = 1  #0.01 
sd_eta     = 1 #0.01 
```

```{r sim_1, ref.label='simulation'}
```

## Decimal error

```{r}
sub=20

# Parameters for random errors
sd_epsilon = 1  #0.01 
sd_eta     = 0.1 #0.01 
```

```{r sim_01, ref.label='simulation'}
```


## Centesimal error

```{r}
sub=20

# Parameters for random errors
sd_epsilon = 1  #0.01 
sd_eta     = 0.01 #0.01 
```

```{r sim_001, ref.label='simulation'}
```


Conclusion: I was adding too much error to the outcome model. Using 0.1 as the standard deviation of the error for the outcome works better.

## 0.5 std deviation

```{r}
sub=20

# Parameters for random errors
sd_epsilon = 1  #0.01 
sd_eta     = 0.5 #0.01 
```

```{r sim_05, ref.label='simulation'}
```

0.5 is not small enough